Data Exploration: IUCN and AquaMaps

### 2015-12-03 data_explore_2015_data.Rmd
figs using 2015 am data

dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
dir_fig <- file.path(getwd(), 'figures')

@@

hist_sm_within_lg <- ggplot(spp_df4, aes(sm_perc, fill = sm_range)) + 
  ggtheme_plot + 
  geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 4, alpha = .8) +
  scale_fill_discrete(name = 'Dataset with smaller range', labels = c('AquaMaps', 'IUCN')) + 
  labs(title = 'Percent of smaller range within larger range',
       x = 'Percent of smaller range that falls within the larger range', 
       y = sprintf('Proportion of all species (n = %s)', nrow(spp_df4)))

@@

### mongo plot time
scatter_arearatio_vs_overlap <- ggplot(spp_list %>% 
                                         transform(popn_category = reorder(popn_category, category_score)),
                                       aes(x = area_ratio, y = sm_perc, 
                                           color = popn_category,
                                           size = popn_category
                                           )) + 
  ggtheme_plot + 
  ### color the quadrant backgrounds:
    annotate("rect", xmin = area_ratio_quartiles[2],  xmax = 100, 
                     ymin = pct_overlap_quartiles[2], ymax = 100, alpha = .3, fill= "green4")  + 
    annotate("rect", xmin = area_ratio_quartiles[2],  xmax =   0, 
                     ymin = pct_overlap_quartiles[2], ymax = 100, alpha = .3, fill= "greenyellow") + 
    annotate("rect", xmin = area_ratio_quartiles[2],  xmax = 100, 
                     ymin = pct_overlap_quartiles[2], ymax =   0, alpha = .3, fill= "orange") + 
    annotate("rect", xmin = area_ratio_quartiles[2],  xmax =   0, 
                     ymin = pct_overlap_quartiles[2], ymax =   0, alpha = .3, fill= "orangered2") + 
  geom_point() + 
  scale_colour_manual(values = c('LC' = 'green4', 
                                 'NT' = 'greenyellow', 
                                 'VU' = 'yellow', 
                                 'EN' = 'orange', 
                                 'CR' = 'orangered2')) +
  geom_vline(x = area_ratio_quartiles[c(1, 3)],
             color = 'grey50', linetype = 'dashed') +
  geom_hline(y = pct_overlap_quartiles[c(1, 3)],
             color = 'grey50', linetype = 'dashed') +
  geom_point(data = spp_list_label, shape = 4, color = 'grey30', size = 4) + 
  geom_text(data = spp_list_label, 
            aes(label = sciname_label), 
            color = 'grey30',
            fontface = 'bold.italic',
            size = 2,
            hjust = .25,
            vjust = 1.5) +
  labs(x = 'Area match: \n(range area smaller : range area larger) * 100%', 
       y = 'Distribution match: \n % overlap of smaller range w/ larger', 
       title = 'Area match vs distribution match',
       color = 'Risk', size = 'Risk')

@@

### Plot the bar chart
barchart_spp_gp_quads <- ggplot(spp_gp_quadrants %>% 
                                  transform(spp_group = reorder(spp_group, pct_q1)), 
                                aes(x = spp_group, fill = quad, weight = pct_quad)) +
  ggtheme_plot + 
  geom_bar(stat = 'bin', alpha = .5) +
  scale_fill_manual(values = c('q1 well-matched'   = 'green4', 
                               'q2 dist-matched'   = 'greenyellow', 
                               'q3 area-matched'   = 'orange', 
                               'q4 poorly matched' = 'orangered2')) +
  geom_text(aes(label = sprintf('n = %s', n_spp), y = .05), hjust = 0, size = 4) +
  coord_flip() +
  labs(x = 'Species Group', 
       y = 'Relative number of species per quadrant', 
       title = 'Species group breakdown by quadrant')

@@

levelplot(am_sp_map, par.settings = myTheme, 
          main = 'Aquamaps Dataset', 
          colorkey = list( at = seq(0, 1063, length = 15)))

levelplot(iucn_sp_map, par.settings = myTheme, 
          main = 'IUCN Dataset', 
          colorkey = list(at = seq(0, 1063, length = 15)))

@@

am_spp_cells_big <- fread(file.path(dir_anx, 'raw/aquamaps_2014/tables/ohi_hcaf_species_native.csv')) %>%
  as.data.frame()
### am_spp_cell_big is missing loiczid info so bring in hcaf to match

hcaf <- fread(file.path(dir_anx, 'raw/aquamaps_2014/tables/hcaf.csv')) %>% #bring this in just to get loiczid
  as.data.frame() %>%
  select(CsquareCode, LOICZID) %>%
  rename(loiczid = LOICZID)

### find total spp to allow for percent calcs later
am_all_n_tot <- am_spp_cells_big %>% 
  select(SpeciesID) %>% 
  unique() %>% nrow()

all_am <- am_spp_cells_big %>%
  select(SpeciesID, CsquareCode) %>%
  group_by(CsquareCode) %>%
  summarize(all_count = n(),
            all_pct   = all_count/am_all_n_tot) %>%
  as.data.frame()

all <- all_am %>%
  left_join(hcaf, by = 'CsquareCode')

all_am_map <- subs(loiczid_raster, all[ , c('loiczid', 'all_count')], 
                   by = 'loiczid', 
                   which = 'all_count', 
                   subsWithNA = TRUE)

levelplot(all_am_map, par.settings = myTheme, main = 'Entire Aquamaps Dataset')
```

![image](figures/all_aquamaps_map.png)

@@

![Absolute change in SPP score](figures/SPP_scenarios/boxplot_abs_diff.png)

![Relative change in SPP score](figures/SPP_scenarios/boxplot_rel_diff.png)

>>>>

dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
dir_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data')
dir_fig <- file.path(getwd(), 'figs2015')

@@

hist_sm_within_lg <- ggplot(spp_df4, aes(sm_perc, fill = sm_range)) + 
  ggtheme_plot + 
  geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 4, alpha = .8) +
  scale_fill_discrete(name = 'Dataset with smaller range', labels = c('AquaMaps', 'IUCN')) + 
  labs(title = 'Percent of smaller range within larger range 2015 data',
       x = 'Percent of smaller range that falls within the larger range', 
       y = sprintf('Proportion of all species (n = %s)', nrow(spp_df4)))

@@

### mongo plot time
scatter_arearatio_vs_overlap <- ggplot(spp_list %>% 
                                         transform(popn_category = reorder(popn_category, category_score)),
                                       aes(x = area_ratio, y = sm_perc, 
                                           color = popn_category,
                                           size = popn_category
                                           )) + 
  ggtheme_plot + 
  ### color the quadrant backgrounds:
    annotate("rect", xmin = area_ratio_quartiles[2],  xmax = 100, 
                     ymin = pct_overlap_quartiles[2], ymax = 100, alpha = .3, fill= "green4")  + 
    annotate("rect", xmin = area_ratio_quartiles[2],  xmax =   0, 
                     ymin = pct_overlap_quartiles[2], ymax = 100, alpha = .3, fill= "greenyellow") + 
    annotate("rect", xmin = area_ratio_quartiles[2],  xmax = 100, 
                     ymin = pct_overlap_quartiles[2], ymax =   0, alpha = .3, fill= "orange") + 
    annotate("rect", xmin = area_ratio_quartiles[2],  xmax =   0, 
                     ymin = pct_overlap_quartiles[2], ymax =   0, alpha = .3, fill= "orangered2") + 
  geom_point() + 
  scale_colour_manual(values = c('LC' = 'green4', 
                                 'NT' = 'greenyellow', 
                                 'VU' = 'yellow', 
                                 'EN' = 'orange', 
                                 'CR' = 'orangered2',
                                 'DD' = 'grey50')) +
  geom_vline(x = area_ratio_quartiles[c(1, 3)],
             color = 'grey50', linetype = 'dashed') +
  geom_hline(y = pct_overlap_quartiles[c(1, 3)],
             color = 'grey50', linetype = 'dashed') +
  geom_point(data = spp_list_label, shape = 4, color = 'grey30', size = 4) + 
  geom_text(data = spp_list_label, 
            aes(label = sciname_label), 
            color = 'grey30',
            fontface = 'bold.italic',
            size = 2,
            hjust = .25,
            vjust = 1.5) +
  labs(x = 'Area match: \n(range area smaller : range area larger) * 100%', 
       y = 'Distribution match: \n % overlap of smaller range w/ larger', 
       title = 'Area match vs distribution match',
       color = 'Risk', size = 'Risk')

@@

### Plot the bar chart
barchart_spp_gp_quads <- ggplot(spp_gp_quadrants %>% 
                                  transform(spp_group = reorder(spp_group, pct_q1)), 
                                aes(x = spp_group, fill = quad, weight = pct_quad)) +
  ggtheme_plot + 
  geom_bar(stat = 'bin', alpha = .5) +
  scale_fill_manual(values = c('q1 well-matched'   = 'green4', 
                               'q2 dist-matched'   = 'greenyellow', 
                               'q3 area-matched'   = 'orange', 
                               'q4 poorly matched' = 'orangered2')) +
  geom_text(aes(label = sprintf('n = %s', n_spp), y = .05), hjust = 0, size = 4) +
  coord_flip() +
  labs(x = 'Species Group', 
       y = 'Relative number of species per quadrant', 
       title = 'Species group breakdown by quadrant 2015 data')

@@

am_lvlplot <- levelplot(am_sp_map, par.settings = myTheme, 
          main = 'Aquamaps Dataset 2015', 
          colorkey = list( at = seq(0, 1063, length = 15)))

png(file.path(dir_fig, 'levelplot_am.png'))
print(am_lvlplot)
dev.off()

iucn_lvlplot <- levelplot(iucn_sp_map, par.settings = myTheme, 
          main = 'IUCN Dataset', 
          colorkey = list(at = seq(0, 1063, length = 15)))

png(file.path(dir_fig, 'levelplot_iucn.png'))
print(iucn_lvlplot)
dev.off()

@@

am_spp_cells_big <- fread(file.path(dir_data, 'aquamaps/v2015/csv/hcaf_sp_native_trunc.csv')) %>%
  as.data.frame()

### find total spp to allow for percent calcs later
am_all_n_tot <- am_spp_cells_big %>% 
  select(speciesid) %>% 
  unique() %>% 
  nrow()

all_am <- am_spp_cells_big %>%
  select(speciesid, loiczid) %>%
  group_by(loiczid) %>%
  summarize(all_count = n(),
            all_pct   = all_count/am_all_n_tot) %>%
  as.data.frame()

all_am_map <- subs(loiczid_raster, all_am[ , c('loiczid', 'all_count')], 
                   by = 'loiczid', 
                   which = 'all_count', 
                   subsWithNA = TRUE)

am_lvlplot_all <- levelplot(all_am_map, par.settings = myTheme, 
                            main = 'Entire Aquamaps Dataset 2015')

png(file.path(dir_fig, 'levelplot_am_all.png'))
print(am_lvlplot_all)
dev.off()

![image](figs2015/all_aquamaps_map.png)

@@

![Absolute change in SPP score](figs2015/SPP_scenarios/boxplot_abs_diff.png)

![Relative change in SPP score](figs2015/SPP_scenarios/boxplot_rel_diff.png)
